---
title: "Xiao_HW5"
name: "Stanley Xiao"
date: "2026-09-20"
format: html
editor: visual
---

# Homework 5

#### Task 1 - Querying an API and a Tidy-Style Function

```{r}
#Install packages as necessary
#install.packages("httr")
#install.packages("jsonlite")
```

```{r}
#Load Packages and API Key
library(httr)
library(jsonlite)
library(tidyverse)

news_api_key <- "74a38983f4844f5a8891b351f068b813"
```

1\. Use GET() from the httr package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the past with a free account.

```{r}
#Query the API using 

news_response <- GET(
  url = "https://newsapi.org/v2/everything",
  query = list(
    q = "artificial intelligence",
    from = Sys.Date() - 14, #14 Days Ago
    language = "en",
    pageSize = 50,
    apiKey = news_api_key
  )
)

news_response
```

2\. Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!

```{r}
#Extract news content by parsing the response into an R object
news_content <- content(news_response, as = "parsed", simplifyVector = TRUE)

#Extract articles from the above content
news_articles <- as_tibble(news_content$articles)

#Look at structure of news_articles
str(news_articles)

#Display first few articles
head(news_articles)

#The resulting object contains info about the articles returned by the query, including the source, author, title, description, URL, publication date, and available article content
```

3\. Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r}
#Query NewsAPI and return the article data frame
query_news <- function(subject, from_date, api_key) {

  response <- GET(
    url = "https://newsapi.org/v2/everything",
    query = list(
      q = subject,
      from = from_date,
      language = "en",
      pageSize = 50,
      apiKey = api_key
    )
  )
  #Parse JSON response
  parsed_response <- content(
    response,
    as = "parsed",
    simplifyVector = TRUE
  )
  as_tibble(parsed_response$articles)
}
```

###### Use your function twice to show it works!

```{r}
#Example 1
travel_news <- query_news(
  subject = "travel",
  from_date = Sys.Date() - 2,
  api_key = news_api_key
)

head(travel_news)

#Example 2
celebrity_news <- query_news(
  subject = "celebrity",
  from_date = Sys.Date() - 1,
  api_key = news_api_key
)

head(celebrity_news)
```

#### Task 2 - Summarizing Roller Coaster Data

Loading roller coaster data

```{r}
coaster_data <- read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")

#Examine variable types and convert to factors as needed
str(coaster_data)
```

Some variables represent categorical variables and are converted to factors below.

```{r}
coaster_data <- coaster_data |>
  mutate(
    Park = as.factor(Park),
    City = as.factor(City),
    State = as.factor(State),
    CensusRegion = as.factor(.data[['Census Region']]),
    AgeGroup = factor(AgeGroup, levels = c(1, 2, 3), labels = c("Older", "Recent", "Newest")),
    Scale = as.factor(Scale),
    Type = as.factor(Type),
    Design = as.factor(Design),
    Inversions = as.factor(.data[['Inversions?']]),
    Make = as.factor(Make)
  )
```

4\. Look at how the data is stored and see if everything makes sense.

```{r}
#Look at structure, dimensions, variable names, and descriptive statistics of the data. Variable types should match its respective type (i.e. chr/factor, int/dbl, and geospatial which won't be used)

str(coaster_data)

dim(coaster_data)

names(coaster_data)

summary(coaster_data)
```

The structure output confirms that categorical variables are stored as factors and quantitative measurements (e.g. top speed, max height, drop, etc.) are stored numerically.

5\. Document any missing values in the data.

```{r}
missing_values <- colSums(is.na(coaster_data))

missing_values[missing_values > 0]
```

TopSpeed has 74 missing values. MaxHeight has 8 missing values. Drop has 128 missing values. Length has 33 missing values. Duration has 66 missing values. Make has 36 missing values.

#### Categorical Variables

6\. Create a one-way contingency table, a two-way contingency table, and a three-way contingency table for some of the factor variables you created previously. Use table() to accomplish this.

-Interpret a number from each resulting table (that is, pick out a value produced and explain what that value means.)

```{r}
#1 Way Table - Coaster Type
table_one_way <- table(Type = coaster_data$Type)
table_one_way

#2 Way Table - Type by Scale
table_two_way <- table(Type = coaster_data$Type, Scale = coaster_data$Scale)
table_two_way

#3 Way Table - Type by Scale by Inversions
table_three_way <- table(Type = coaster_data$Type, Scale = coaster_data$Scale, Inversions = coaster_data$Inversions)
table_three_way
```

**One-way table interpretation:** There are 530 steel roller coasters in the dataset.

**Two-way table interpretation:** There are 216 roller coasters that are both steel and classified as Extreme.

**Three-way table interpretation:** There are 166 roller coasters that are steel, classified as Extreme, and have inversions.

7\. Create a conditional two-way table using table(). That is, condition on one variable’s setting and create a two-way table. Do this using two different methods:

\- Once, by subsetting the data (say with filter()) and then creating the two-way table

\- Once, by creating a three-way table and subsetting it

```{r}
#Method 1 -- Subsetting the Data only in the Midwest Region
midwest_coasters <- coaster_data |>
  filter(CensusRegion == "Midwest")

###Create a 2-way table for the filtered obs
midwest_table_method1 <- table(
  midwest_coasters$Scale,
  midwest_coasters$Type
)

midwest_table_method1

#Method 2 - Subsetting a 3-way table
###Obtaining the same result via a 3 way contingency table
conditional_table <- table(
  coaster_data$Scale,
  coaster_data$Type,
  coaster_data$CensusRegion
)

midwest_table_method2 <- conditional_table[, , "Midwest"]
midwest_table_method2
```

8\. Create a two-way contingency table using group_by() and summarize() from dplyr. Then use pivot_wider() to make the result more like the output from table().

```{r}
#Count each combination of Scale & Type
scale_type_summary <- coaster_data |>
  group_by(Scale, Type) |>
  summarize(
    count = n(),
    .groups = "drop"
  )

scale_type_wide <- scale_type_summary |>
  pivot_wider(
    names_from = Type,
    values_from = count,
    values_fill = 0
  )

scale_type_wide
```

9\. Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for the plots.

```{r}
#Stacked Bar Graph of Scale by Type
ggplot(
  coaster_data,
  aes(x = Scale, fill = Type)
) +
  geom_bar() +
  labs(
    title = "Stacked Bar Graph of Roller Coaster Type by Scale",
    x = "Coaster Scale",
    y = "Number of Roller Coasters",
    fill = "Coaster Type"
  )

#Side by side bar graph
ggplot(
  coaster_data,
  aes(x = Scale, fill = Type)
) +
  geom_bar(position = "dodge") +
  labs(
    title = "Side-by-Side Bar Graph of Roller Coaster Type by Scale",
    x = "Coaster Scale",
    y = "Number of Roller Coasters",
    fill = "Coaster Type"
  )
```
