---
title: "sxiao3_HW2_Individual"
format: html
editor: visual
---

## Task 1: Basic Vector Practice

1.  Assigning names for pre-treatment BP values, post-treatment BP values

```{r}
#Pre-Treatment BP values
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

#Post-Treatment BP values
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2.  Assigning Subject Names for each Pre/Post-treatment BP values

```{r}
names(pre_bp) <- paste("Subject", 1:20, sep="_")
names(post_bp) <- paste("Subject", 1:20, sep="_")
```

3.  Change in BP for each Patient

```{r}
diff_bp <- pre_bp - post_bp
```

4.  Average Decrease in BP Across All Patients

```{r}
avg_change <- mean(diff_bp)
```

5.  Which Patients Experienced Decrease in BP After Treatment (Positive Change)

```{r}
lower_bp <- which(diff_bp > 0)
```

6.  Subset Vector in \[HW2 1.5\] to Only Return Positive Change

```{r}
only_lower <- diff_bp[lower_bp]
```

7.  Average Decrease in BP for Positive Change (BP Decrease)

```{r}
avg_lower <- mean(only_lower)
```

## Task 2: Basic Data Frame Practice

1.  Create a data frame object with four columns corresponding to your data above: patient, pre_bp, post_bp, and diff_bp

```{r}
bp <- data.frame(patient = names(pre_bp), pre_bp, post_bp, diff_bp, row.names=NULL)

bp
```

2.  Return only rows where the diff_bp column is negative. (Use \[ or learn about the subset() function if you’d like. If you use \[, don’t reference the original vector from the first part, access the column of the data frame when looking at making a comparison.

```{r}
subset(bp, diff_bp < 0)
```

3.  Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120. Recall you can use \$ to reference a column. If you reference a column that doesn’t exist, and save a vector (of appropriate length in it), that vector becomes a column of your data frame! Similar to the previous question, don’t reference the original vector from the first part, access the column of the data frame when looking at making a comparison.

```{r}
bp$bp_post_lt_120 <- bp$post_bp < 120
```

4.  Finally, print the data frame out nicely in your final document by modifying the code below appropriately.

```{r}
knitr::kable(bp)
```

## Task 3: List Practice

1.  Create a new data frame with this data in it that is similar to the data frame from task 2 (including the new column).

```{r}
#Now we're adding in another experiment where the 'treatment' was actually a placebo.
placebo_bp_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

placebo_bp_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

placebo <- data.frame(patient=paste('Subject', 1:10, sep="_"), pre_bp = placebo_bp_pre, post_bp = placebo_bp_post, diff_bp = placebo_bp_pre - placebo_bp_post, bp_post_lt_120 = placebo_bp_post < 120)
```

2.  Now create and store a list with two elements: • 1st element named treatment and contains the first data frame you created. • 2nd element named placebo and contains the second data frame you created.

```{r}
bp_list <- list(treatment = bp, placebo = placebo)
```

3.  Access the first list element using three different types of syntax.

```{r}
# Method 1
bp_list$treatment
# Method 2
bp_list[["treatment"]]
# Method 3
bp_list[[1]]
```

4.  In one line, access the placebo data frame, pre_bp column.

```{r}
bp_list$placebo$pre_bp
```

## Task 4: Control Flow Practice

1.  Suppose we want to characterize the post-treatment (or placebo) blood pressure measurement as optimal (≤ 120), borderline (120 \< bp ≤ 130), and high (\> 130). First, create a new column in each data frame from above called status. You can do this via
    1.  your_df\$status \<- character(20) #or 10 depending on number of observations

```{r}
bp_list$treatment$status <- character(20)
bp_list$placebo$status <- character(10)
```

2.  For the non-placebo data frame (within the list), create a for loop and use if/then/else logic to create the status column’s values.

```{r}
for (i in 1:nrow(bp_list$treatment)) {
  if (bp_list$treatment$bp_post[i] <= 120) {
    bp_list$treatment$status[i] <- "Optimal"
  } else if (bp_list$treatment$bp_post[i] <= 130) {
    bp_list$treatment$status[i] <- "Borderline"
  } else {
    bp_list$treatment$status[i] <- "High"
  }
}
```

3.  Repeat for the placebo data frame (within the list).

```{r}
for (i in 1:nrow(bp_list$placebo)) {
  if (bp_list$placebo$bp_post[i] <=  120) {
    bp_list$placebo$status[i] <- "Optimal"
  } else if (bp_list$placebo$bp_post[i] <= 130) {
    bp_list$placebo$status[i] <- "Borderline"
  } else {
    bp_list$placebo$status[i] <- "High"
  }
}
```
